---
title: "SEA 2018 report"
output:
  pdf_document:
    keep_tex: yes
    number_sections: yes
    toc: yes
  html_document:
    df_print: tibble
    number_sections: yes
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r setup-libraries}
rm(list = ls())
library(tidyverse)
library(knitr)

qup <- function(x) quantile(x, 3/4)
qdown <- function(x) quantile(x, 1/4)
```

# Experimental setup
## Time measurements
We run a program $k$ times with and without the optimization and recorded the sum construction time of the MS and RUNS vectors. The plots report (median, with quartile ranges) of the speedup of each optimized time $t_i^{\mathtt{opt}}$ relative to the average non-optimized time in the construction time of the MS vector. In other words  
$$
d^{(i)} = \frac{\bar{t}_{\mathtt{non\_opt}}}{t^{(i)}_{\mathtt{opt}}}
$$
with $\bar{t}_{\mathtt{non\_opt}} = 1/n \sum t^{(i)}_{\mathtt{non\_opt}}$, and $i = 1, \ldots, k$. 

The boxplots report the raw times.

# WL tests
## Input data
We perform tests on the Weiner Link optimizations on 4 types of input. 

 * Index string with repeats, query string random (code: `rep_dis`)
 * Index string with repeats, query string similar to index (code: `rep_sim`)
 * Index string random, query string random (code: `rnd_dis`)
 * Index string random, query string similar to index (code: `rnd_sim`)

Further, we generate all of the above input data for two alphabet sizes: $\Sigma_1| = 4$ and  $\Sigma_2| = 20$. For all input types, the index string is of length 100MB and the query 500KB.
```{r wl_data, echo=FALSE}
read_many_csv <- function(trials){
  do.call(rbind, lapply(trials, function(trial) {
    read_csv(sprintf("../wl_tests/sea18/trial_%d/rank_timing.csv", trial)) %>%
      mutate(ntrial = trial)
  }))
}

time_data <- (
  read_many_csv(seq(1, 10)) %>% 
    filter(item %in% c("ms_bvector", "runs_bvector")) %>% 
    separate(b_path, into=c('stype', 'slen', 'ttype', 'tlen', 'alp')) %>% 
    separate(label, into=c("lazy", "fail", "drank", "maxrep")) %>% 
    unite(col = inp_type, stype, ttype, sep = "_") %>% 
    mutate(lazy = ifelse(lazy == "l1", "lazy", "nonlazy"), 
           fail = ifelse(fail == "f1", "fail", "nonfail"),
           drank = ifelse(drank == "r0", "srank", "drank"),
           maxrep = ifelse(maxrep == "m0", "nonmaxrep", 
                           ifelse(maxrep == "m1", 
                                  "maxrep_vanilla", "maxrep_rc"))) %>%
    select(ntrial, alp, item, lazy, fail, drank, maxrep, inp_type, value)
)

time_data <- (
  time_data %>% 
    spread(item, value) %>% 
    mutate(bvector = ms_bvector + runs_bvector, value = bvector) %>% 
    select(-ms_bvector, -runs_bvector)
)

prange_plot <- function(dt){
  ggplot(dt, aes(inp_type, value)) +
    geom_pointrange(stat = "summary", fatten = 2, 
                    fun.y = median, fun.ymin = qdown, fun.ymax = qup) +
    geom_jitter(alpha=0.2, width = 0.2) +
    geom_hline(yintercept = 1.0, color='blue') + theme_bw() +
    facet_grid(~alp) +
    ylab("Speedup") + xlab("Input type")
}
```


## Double rank versus single rank
```{r drank_plot_better}
opt_dt <- (time_data %>% 
             filter(lazy == "nonlazy", fail == "nonfail",
                    drank == "drank", maxrep == "nonmaxrep")
)
baseline_dt <- (time_data %>% 
                  filter(lazy == "nonlazy", fail == "nonfail",
                  drank == "srank", maxrep == "nonmaxrep")
)
dt <- (opt_dt %>% 
         left_join(baseline_dt, by = c("ntrial", "alp", "inp_type")) %>%
         mutate(value = bvector.y / bvector.x) %>% 
         select(ntrial, alp, inp_type, value))
prange_plot(dt) 
```

## Lazy versus nonlazy
```{r lazy_plot}
opt_dt <- (time_data %>% 
             filter(lazy == "lazy", fail == "nonfail",
                    drank == "srank", maxrep == "nonmaxrep")
)
baseline_dt <- (time_data %>% 
                  filter(lazy == "nonlazy", fail == "nonfail",
                  drank == "srank", maxrep == "nonmaxrep")
)
dt <- (opt_dt %>% 
         left_join(baseline_dt, by = c("ntrial", "alp", "inp_type")) %>%
         mutate(value = bvector.y / bvector.x) %>% 
         select(ntrial, alp, inp_type, value))
prange_plot(dt)
```


## Double rank and fail versus double rank
```{r fail_plot}
opt_dt <- (time_data %>% 
             filter(lazy == "nonlazy", fail == "fail",
                    drank == "drank", maxrep == "nonmaxrep")
)
baseline_dt <- (time_data %>% 
                  filter(lazy == "nonlazy", fail == "nonfail",
                  drank == "drank", maxrep == "nonmaxrep")
)
dt <- (opt_dt %>% 
         left_join(baseline_dt, by = c("ntrial", "alp", "inp_type")) %>%
         mutate(value = bvector.y / bvector.x) %>% 
         select(ntrial, alp, inp_type, value))
prange_plot(dt)
```


## Double rank, Lazy, and Rank and fail versus Single rank, nonlazy, and nonfail
```{r allflags_plot}
opt_dt <- (time_data %>% 
             filter(lazy == "lazy", fail == "fail",
                    drank == "drank", maxrep == "nonmaxrep")
)
baseline_dt <- (time_data %>% 
                  filter(lazy == "nonlazy", fail == "nonfail",
                  drank == "srank", maxrep == "nonmaxrep")
)
dt <- (opt_dt %>% 
         left_join(baseline_dt, by = c("ntrial", "alp", "inp_type")) %>%
         mutate(value = bvector.y / bvector.x) %>% 
         select(ntrial, alp, inp_type, value))

prange_plot(dt)
```


## Maxrep
### Vanilla Maxrep vs. non-maxrep
```{r maxrep1_plot}
opt_dt <- (time_data %>% 
             filter(lazy == "nonlazy", fail == "fail",
                    drank == "drank", maxrep == "maxrep_vanilla")
)
baseline_dt <- (time_data %>% 
                  filter(lazy == "nonlazy", fail == "fail",
                  drank == "drank", maxrep == "nonmaxrep")
)
dt <- (opt_dt %>% 
         left_join(baseline_dt, by = c("ntrial", "alp", "inp_type")) %>%
         mutate(value = bvector.y / bvector.x) %>% 
         select(ntrial, alp, inp_type, value))

prange_plot(dt)
```

### Rank&check Maxrep versus Vanilla Maxrep
```{r maxrep2_plot}
opt_dt <- (time_data %>% 
             filter(lazy == "nonlazy", fail == "fail",
                    drank == "drank", maxrep == "maxrep_rc")
)
baseline_dt <- (time_data %>% 
                  filter(lazy == "nonlazy", fail == "fail",
                  drank == "drank", maxrep == "maxrep_vanilla")
)
dt <- (opt_dt %>% 
         left_join(baseline_dt, by = c("ntrial", "alp", "inp_type")) %>%
         mutate(value = bvector.y / bvector.x) %>% 
         select(ntrial, alp, inp_type, value))

prange_plot(dt)
```


# Optimizations on parent operations
## Input data
We generate the index input string with repetitions as follows. We generate a random seed block $b$ of length 200. Next, we generate blocks of the same length $b_k$ by introducing $k$ mutations on $b$. The index string of length 10MB is $b\circ b_k^{(1)} \ldots \circ b_k^{(4999)}$.  

The query string is obtained as a concatenation of labels from nodes of the suffix tree of $s$. We select nodes with node depth of at least 10 and string length at most 170 for a total string length of 103KB. We separate the labels with a sentinel character that does not appear in $s$. 

Furthermore, we perform experiments for various choices of $1 \geq k \geq |b|$.

The plot below shows a histogram of the length of consecutive parent operations. This quantity is important since the speedup of this optimization is proportional to the length of sequence of parent operations. Importantly, the optimization might not even be beneficial if the length of the sequence of parent operations is less than 3.

```{r}
bdir <- '../input_stats_data/exploration/index_based_query/rev_labels/'
istat_data <- (
  do.call(rbind, 
          lapply(list.files(bdir, pattern = '*.istat.csv', full.names = TRUE), 
                 function(p){
                   read_csv(p) %>% mutate(b_path = basename(p))
                 })
          ) %>% 
    filter(measuring == 'pseq') %>% 
    mutate(key = as.numeric(key)) %>% 
    separate(b_path, c('rep', 'k', 's1', 's2')) %>% 
    select(-s1, -s2, -rep) %>% 
    mutate(k = factor(as.numeric(k)))
)
ggplot(istat_data, aes(key, value, color=where)) + 
  geom_line() + 
  scale_colour_brewer(palette = "Dark2") + theme_bw() +
  facet_wrap(~k) + xlim(0, 100) + 
  xlab("Length of parent sequence") + ylab("Count")
```

## LCA versus parent sequence
```{r}
read_many_csv <- function(trials){
  do.call(rbind, lapply(trials, function(trial) {
    trial_dir <- sprintf("../parent_tests/sea18/trial_%d", trial)
    do.call(rbind, 
            lapply(list.files(trial_dir, pattern = "*.csv", full.names = TRUE), 
                   function(p) read_csv(p) %>% mutate(ntrial = trial)))
    }))
}

time_data <- (read_many_csv(seq(1, 10)) %>% 
                filter(item %in% c('runs_bvector', "ms_bvector")) %>%
                separate(b_path, c('rep', 'k')) %>% 
                select(-len_t, -len_s, -rep) %>% 
                spread(label, value) %>% 
                group_by(ntrial, k) %>% 
                summarise(lca = sum(lca), pseq = sum(pseq)) %>% 
                mutate(value = pseq / lca, k = factor(as.numeric(k)))
)
time_data

prange_plot <- function(dt){
  ggplot(dt, aes(k, value)) +
    geom_pointrange(stat = "summary", fatten = 2, 
                    fun.y = median, fun.ymin = qdown, fun.ymax = qup) +
    geom_jitter(alpha=0.2, width = 0.1) +
    geom_hline(yintercept = 1.0, color='blue') + theme_bw() + 
    ylab("Speedup") + xlab("Block mutations")
}
prange_plot(time_data)
```

#Parallelization
```{r}
rm(list=ls())
library(tidyverse)
dd <- (read_csv('../../fast_ms/parallelization_bug/stats.csv') 
 %>% separate(fname, c('inp_type', 'nslices', 'nthreads', 'suff'))
 %>% mutate(nthreads = as.integer(nthreads))
 %>% filter(item != 'answer',
            item != 'runs_cst',
            item != 'ms_cst',
            item != 'ms_merge',
            item != 'runs_merge')
 %>% separate(item, c('section', 'action'))
 %>% group_by(inp_type, nslices, nthreads, section, action) 
 %>% summarise(avg = mean(value)
               #mid = median(value), 
               #up = qup(value), down = qdown(value)
               )
  %>% ungroup()
)
dd

ggplot(dd
       %>% filter(action %in% c('total'), inp_type == 'protein')
       %>% spread(action, avg),
       aes(nthreads, total, color=section)) +
  geom_point() + geom_line() +
  facet_grid(nslices~inp_type)


ggplot(dd %>% filter(action %in% c('correct', 'vec')), aes(nthreads, avg, color=action)) + geom_smooth(se = FALSE, method=lm) + geom_point() + facet_wrap(~inp_type, scales = 'free_y')

ggplot(dd %>% filter(section != 'comp', action != 'vec'), aes(nthreads, avg, color=action)) + geom_line() + geom_point() + facet_grid(section~inp_type, scales = 'free_x')

get_theoretical <- function(xx){
  nthreads = seq(1, 15)
  vals <- sapply(nthreads, function(n) xx$avg / n)
  rbind(
    tibble(avg = vals[1,], inp_type = xx$inp_type[1], nthreads = nthreads),
    tibble(avg = vals[2,], inp_type = xx$inp_type[2], nthreads = nthreads)
  ) %>% mutate(section = xx$section[1], action = xx$action[1])
}

ee <- do.call(
  rbind,
  lapply(as.numeric((dd %>% group_by(nslices) %>% summarise(n=n()))$nslices),
         function(ns){
           xx <- (dd %>% filter(nslices == ns, section == 'comp', nthreads == 1))
           (get_theoretical(xx) %>% mutate(nslices = ns))
           }
         )
  )
(dd %>% filter(nslices == 10, section == 'comp', nthreads == 1))
ee %>% group_by(inp_type) %>% summarise(n=n())
dd %>% filter(section == 'comp')

ggplot(
  rbind(dd %>% filter(section == 'comp') %>% mutate(experiment = 'actual'),
        ee %>% mutate(experiment = 'theoretical')) %>% select(-section, -action),
  aes(nthreads, avg, color=experiment)) + geom_line() +
  facet_wrap(inp_type~nslices, scales = 'free_y')

ggplot(dd %>% filter(nslices == 1000, section == 'comp'), aes(nthreads, avg)) + geom_line() + geom_point() + facet_wrap(~inp_type, scales = 'free_x')
```



```{r}
rm(list=ls())
dd <- (read_csv('../parallelization_data/stats.csv') 
 %>% separate(fname, c('stype', 'slen', 'ttype', 'tlen', 'alp', 'nthreads', 'suff'))
 %>% unite(col = inp_type, stype, ttype, sep='_')
 %>% select(-slen, -tlen, -suff) %>% mutate(nthreads = as.integer(nthreads))
 %>% filter(item != 'answer', item != 'runs_cst', item != 'ms_cst', item != 'ms_merge', item != 'runs_merge')
 %>% separate(item, c('section', 'action'))
 %>% select(-alp, -len_s, -len_t)
 %>% group_by(inp_type, nthreads, section, action) 
 %>% summarise(avg = mean(value)
               #mid = median(value), 
               #up = qup(value), down = qdown(value)
               )
  %>% ungroup()
)
dd

ggplot(dd %>% filter(action %in% c('correct', 'build')) %>% spread(action, avg), aes(correct, build)) + geom_point() + geom_smooth() + facet_wrap(~inp_type)


ggplot(dd %>% filter(action %in% c('correct', 'build')), aes(nthreads, avg, color=action)) + geom_smooth(se = FALSE, method=lm) + geom_point() + facet_wrap(~inp_type, scales = 'free_y')

ggplot(dd %>% filter(section != 'comp', action != 'vec'), aes(nthreads, avg, color=action)) + geom_line() + geom_point() + facet_grid(section~inp_type, scales = 'free_x')

ggplot(dd %>% filter(section == 'comp'), aes(nthreads, avg)) + geom_line() + geom_point() + facet_wrap(~inp_type, scales = 'free_x')
```

# Genome tests
## Figure 4
```{r fig4}
prange_plot <- function(dt){
  ggplot(dt, aes(flag, value)) +
    geom_jitter(alpha=0.2, width = 0.3) +
    geom_pointrange(stat = "summary", fatten = 2, 
                    fun.y = median, fun.ymin = qdown, fun.ymax = qup) +
    geom_hline(yintercept = 1.0, color='blue') +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1), axis.title.x = element_blank(),
          legend.position = "none") + 
    ylab("Speedup") + scale_color_manual(values=c("#000000", "#d95f02"))
}



flags <- c("F_DR_M", "F_DR_MC", "F_DR_LZ", "DR_LZ", "LZ", "F_DR", "DR", "F_DR_LZ_P", "DR_LZ_P", "LZ_P", "F_DR_P", "DR_P", "P")
pdt <- rbind(read_csv('figures4and5_data/figure4-proteome.txt') %>% 
              mutate(setup = "Experiment1"), 
            read_csv('figures4and5_data/figure4-proteome-bacteria.txt') %>% 
              mutate(setup = "Experiment2")
            )

gdt <- rbind(read_csv('figures4and5_data/figure4-genome.txt') %>% 
              mutate(setup = "Experiment1"), 
            read_csv('figures4and5_data/figure4-genome-bacteria.txt') %>% 
              mutate(setup = "Experiment2")
            )
dt <- (rbind(gdt %>% mutate(sequence = "Genome"), pdt %>% mutate(sequence = "Proteome")) %>% 
         mutate(flag = factor(flag, levels = flags, ordered = TRUE)))

prange_plot(dt %>% filter(setup=="Experiment1")) + geom_point(data=dt %>% filter(setup=="Experiment2"), aes(flag, value), color="#d95f02") + facet_grid(~sequence) 
```

## Figure 5
```{r fig5}
prange_plot <- function(dt){
  ggplot(dt, aes(setup, value), color="#000000") +
    geom_pointrange(stat = "summary", fatten = 2, 
                    fun.y = median, fun.ymin = qdown, fun.ymax = qup) +
    geom_jitter(alpha=0.2, width = 0.1) +
    geom_hline(yintercept = 1.0, color='blue') +
    theme_bw() +
    theme(legend.position = "none", axis.title.x = element_blank()) +
    ylab("Competitor / This paper") + scale_x_discrete(labels = function(x) str_wrap(x, width = 10))
}

mem_dt <- rbind(read_csv('figures4and5_data/figure5-memory-genome-bacteria.txt'),
                read_csv('figures4and5_data/figure5-memory-genome-similar.txt'),
                read_csv('figures4and5_data/figure5-memory-genome.txt'),
                read_csv('figures4and5_data/figure5-memory-proteome-bacteria.txt'),
                read_csv('figures4and5_data/figure5-memory-proteome-similar.txt'),
                read_csv('figures4and5_data/figure5-memory-proteome.txt')
                )

tim_dt <- rbind(read_csv('figures4and5_data/figure5-time-genome-bacteria.txt'),
                read_csv('figures4and5_data/figure5-time-genome-similar.txt'),
                read_csv('figures4and5_data/figure5-time-genome.txt'),
                read_csv('figures4and5_data/figure5-time-proteome-bacteria.txt'),
                read_csv('figures4and5_data/figure5-time-proteome-similar.txt'),
                read_csv('figures4and5_data/figure5-time-proteome.txt')
                )
dt <- (rbind(mem_dt %>% mutate(measuring = "Memory"), tim_dt %>% mutate(measuring = "Time")) %>% 
         mutate(measuring = factor(measuring, levels=c("Time", "Memory"), ordered = TRUE)))

prange_plot(dt %>% filter(sequence=="non-bacteria")) + geom_point(data=dt %>% filter(sequence=="bacteria"), aes(setup, value), color="#d95f02") + facet_wrap(~measuring, scales="free_y")
```


## parallel on real data
```{r genome-tests, eval=FALSE}
library(tidyverse)
rm(list=ls())
n_threads_original <- c(1, 2, 4, 8, 16, 24, 32, 40, 48)
cbbPalette <- c("#009E73", "#D55E00", "#0072B2", "#000000")

df_i <- read_csv('./parallel_real_data/individual_data.csv')
df_s <- (read_csv('./parallel_real_data/similar_data.csv')
         %>% separate(name, sep='-', into=c('inp_type', 'name')))
df <- (rbind(df_i, df_s)
       %>% filter(name %in% c('runs_build', 'runs_correct', 'runs_total',
                              'ms_build', 'comp_total'))
       %>% mutate(nthreads = n_threads_original[nthreads])
       %>% mutate(inp_type = factor(inp_type, 
                                    levels = c('individ-different',
                                               'genome', 
                                               'proteome',
                                               'individ-same'), ordered = TRUE))
)

# first column
df1 <- (df %>% filter(nthreads > 1, name == 'runs_build')
       %>% left_join((df
                      %>% filter(nthreads == 1)
                      %>% select(-nthreads)
                      %>% rename(thr1_time_ms = time_ms)),
                     by = c('name', 'inp_type', 'run'))
       %>% mutate(time_ratio = time_ms / thr1_time_ms,
                  ideal_time_ratio = 1 / nthreads,
                  tit = 'Construction of runs')
)

p1 <- ggplot(df1, aes(nthreads, log2(time_ratio), color = inp_type)) +
  geom_jitter(width = 0.05, shape=1) +
  geom_line(aes(nthreads, log2(ideal_time_ratio)), color='black') +
  scale_x_continuous(trans = 'log2', breaks = n_threads_original, name = "") +
  scale_y_continuous(name = "Time ratio (log2)", breaks = seq(-1, -6, -1.5)) +
  scale_color_manual(values = cbbPalette) +
  theme_bw() + theme(legend.position = "none")
p1

q1 <- ggplot(df1 %>% filter(nthreads > 1),
             aes(nthreads, (time_ratio - ideal_time_ratio) / ideal_time_ratio, color = inp_type)) +
  geom_jitter(width = 0.05, shape=1) + 
  scale_x_continuous(trans = 'log2', breaks = n_threads_original, name = "N. threads") +
  scale_y_continuous(name = "Offset") +
  scale_color_manual(values = cbbPalette) +
  theme_bw() + theme(legend.position = "none")



# second column
df2 <- (
  inner_join(df %>% filter(name == 'runs_correct'),
             (df %>% filter(name == 'runs_total')
              %>% select(-name) %>% rename(tot_time_ms = time_ms)),
             by = c('run', 'nthreads', 'inp_type'))
  %>% mutate(time_ratio = time_ms / tot_time_ms)
  %>% filter(nthreads > 1)
  %>% mutate(inp_type = factor(inp_type, 
                               levels = c('individ-different',
                                          'genome', 'proteome',
                                          'individ-same'), ordered = TRUE))
)
df2 

p2 <- ggplot(df2 %>% mutate(tit = 'Correction of runs'),
             aes(nthreads, time_ratio, color = inp_type)) +
  geom_jitter(width = 0.05, shape=1) + #theme(legend.position = "none") +
  scale_x_continuous(trans = 'log2', breaks = n_threads_original, name = "") + 
  scale_y_continuous(name = "Fraction of total time for runs") +
  scale_color_manual(values = cbbPalette) +
  theme_bw() + theme(legend.position = "none")


q2 <- ggplot(df2 %>% filter(inp_type != 'individ-same') %>% mutate(tit = 'Correction runs (zoom)'),
             aes(nthreads, time_ratio, color = inp_type)) +
  geom_jitter(width = 0.05, shape=1) +
  theme(legend.position = "none") +
  scale_x_continuous(trans = 'log2', breaks = n_threads_original, name = "N. threads") +
  scale_y_continuous(name = "Fraction of total time for runs") +
  scale_color_manual(values = cbbPalette) +
  theme_bw() + theme(legend.position = "none")



# third column
df3 <- (df %>% filter(nthreads > 1, name == 'ms_build')
       %>% left_join((df
                      %>% filter(nthreads == 1)
                      %>% select(-nthreads)
                      %>% rename(thr1_time_ms = time_ms)),
                     by = c('name', 'inp_type', 'run'))
       %>% mutate(time_ratio = time_ms / thr1_time_ms,
                  ideal_time_ratio = 1 / nthreads,
                  tit = 'Construction of ms')
)
p3 <- ggplot(df3, aes(nthreads, log2(time_ratio), color = inp_type)) +
  geom_jitter(width = 0.05, shape=1) +
  geom_line(aes(nthreads, log2(ideal_time_ratio)), color='black') +
  scale_x_continuous(trans = 'log2', breaks = n_threads_original, name = "") +
  scale_y_continuous(name = "Time ratio (log2)", breaks = seq(0, -6, -1.5)) +
  scale_color_manual(values = cbbPalette) +
  theme_bw() + theme(legend.position = "none")

q3 <- ggplot(df3 %>% filter(nthreads > 1, inp_type != 'individ-same'),
             aes(nthreads, (time_ratio - ideal_time_ratio) / ideal_time_ratio, color = inp_type)) +
  geom_jitter(width = 0.05, shape=1) + 
  scale_x_continuous(trans = 'log2', breaks = n_threads_original, name = "N. threads") +
  scale_y_continuous(name = "Offset") +
  scale_color_manual(values = cbbPalette) +
  theme_bw() + theme(legend.position = "none")

df4 <- (df %>% filter(nthreads > 1, name == 'comp_total')
       %>% left_join((df
                      %>% filter(nthreads == 1)
                      %>% select(-nthreads)
                      %>% rename(thr1_time_ms = time_ms)),
                     by = c('name', 'inp_type', 'run'))
       %>% mutate(time_ratio = time_ms / thr1_time_ms,
                  ideal_time_ratio = 1 / nthreads,
                  tit = 'Total')
)
p4 <- ggplot(df4, aes(nthreads, log2(time_ratio), color = inp_type)) +
  geom_jitter(width = 0.05, shape=1) +
  geom_line(aes(nthreads, log2(ideal_time_ratio)), color='black') +
  scale_x_continuous(trans = 'log2', breaks = n_threads_original, name = "") +
  scale_y_continuous(name = "Time ratio (log2)", breaks = seq(0, -6, -1.5)) +
  scale_color_manual(values = cbbPalette) +
  theme_bw() + theme(legend.position = "none")

q4 <- ggplot(df4 %>% filter(inp_type != 'individ-same'),
             aes(nthreads, (time_ratio - ideal_time_ratio) / ideal_time_ratio, color = inp_type)) +
  geom_jitter(width = 0.05, shape=1) + 
  scale_x_continuous(trans = 'log2', breaks = n_threads_original, name = "N. threads") +
  scale_y_continuous(name = "Offset") +
  scale_color_manual(values = cbbPalette) +
  theme_bw() + theme(legend.position = "none")


# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- lapply(c(list(...), plotlist), function(p) p + facet_wrap(~tit))
  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
multiplot(p1, q1, p2, q2, p3, q3, p4, q4, cols=4)
#layout = matrix(c(1, 2, 1, 2, 1, 2, 3, 4, 3, 4, 3, 4, 5, 6, 5, 6, 5, 6, 7, 8, 7, 8, 7, 8, 7, 8), nrow=2))

```


## Range queries
```{r}
library(tidyverse)
rm(list=ls())
block_sizes <- c(0, 2^seq(2, 12))
block_sizes <- c(0, 4, 128, 512, 1024, 4096)

dd <- do.call(rbind, lapply(block_sizes,
       function(i) {
         tmpl <- '../range_tests/m.t.ms.%d.csv'
         path <- sprintf(tmpl, i)
         read_csv(path) %>% mutate(block_size = i)
       }
) )  %>% group_by(block_size) %>% mutate(index = rep(seq(1, n() / 6), each=6)) %>% ungroup()

ggplot(dd %>% filter(method == 'algorithm'), aes(index, time_ms, color=factor(block_size))) + geom_line() + geom_point()
# remove trivial 
#ggplot(dd %>% filter(method == 'algorithm', block_size > 0), aes(index, time_ms, color=factor(block_size))) + geom_line() + geom_point()

ggplot(dd %>% filter(method == 'comp_total'), aes(index, time_ms, color=factor(block_size))) + geom_line() + geom_point()

ggplot(dd %>% filter(!(method %in% c("computation", "comp_total")), block_size > 0), aes(method, time_ms, color=factor(block_size))) + geom_boxplot()

dd %>% filter(method == 'algorithm') %>% group_by(block_size) %>% summarise(time_ms = mean(time_ms))

mdd <- (
  dd %>% filter(method == 'algorithm') 
  %>% group_by(block_size, index) %>% summarise(time_ms=mean(time_ms)) 
  %>% ungroup() 
  %>% inner_join(read_csv('../range_tests/memory_usage.csv'), 
                 by = c("block_size"))
) %>% filter(block_size > 0)

ggplot(mdd, aes(factor(round(peak / (1024^2), 0)), time_ms / 1000, color=factor(block_size))) + geom_boxplot() + geom_jitter(width = 0.1) +
  labs(x = "Space (MB)", y = "Time (s)")
```

